---
title: "Youth unemployment and turnout in Spain"
author: "Rafael Goldzweig & Rita Alvarez"
date: "December 2, 2016"
output:
  pdf_document:
    number_sections: yes
    toc: yes
bibliography: References.bib
---

```{r options, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(message = FALSE)
# opts_chunk$set(tidy=FALSE)
# opts_chunk$set(fig.path='figs/', cache.path='cache/graphics-', 
#         fig.align='center', fig.width = 5, fig.height = 5, fig.show = 'hold',
#         cache = TRUE, par = TRUE)
```

```{r library, message=FALSE, echo=FALSE, include=FALSE}
library(pxR)
library(MicroDatosEs)
library(XML)
library(zoo)
library(rvest)
library(dplyr)
library(magrittr)
library(foreign)
library(readxl)
install.packages("xlsx", repos = 'http://cran.us.r-project.org')
library(xlsx)
library(ggplot2)
library(pastecs)
library(stargazer)
library(gplots)
```

```{r setup, message=FALSE, echo=FALSE, include=FALSE}
# Clear Global environment
rm(list=ls())

# set WD
#setwd('C:/Users/Rita/Desktop/Final Data Project')
```

```{r turnout data, results='asis', tidy=TRUE, echo=FALSE}
# Read Turnout data

Turnout2004_2008_2011_2015 <- read_excel("C:/Users/Rita/Desktop/Assignment-3/ElectionsResult_2015_2011_2008_2004.xlsx", col_names = TRUE)
```

```{r unemployment data, results='asis', tidy=TRUE, echo=FALSE}
# Read Unemployment data

YouthUnemployment2004_2015 <- read_excel("C:/Users/Rita/Desktop/Assignment-3/YouthUnemployment2015_2011_2008_2004.xlsx", col_names = TRUE)
```

```{r controls, results='asis', tidy=TRUE, echo=FALSE}
# Read control variables data

Controlvariables <- read_excel("C:/Users/Rita/Desktop/Assignment-3/Controlvariables.xlsx", col_names = TRUE)

```

```{r dataset1, results='asis', tidy=TRUE, echo=FALSE}
# Merge the dataset of electoral results with the dataset of unemployment rates. Change columns as numeric and clean one column

Spain_data_1 <- merge(Turnout2004_2008_2011_2015, YouthUnemployment2004_2015, by=c("Region","Year"), all=TRUE)
```

```{r dataset2, results='asis', tidy=TRUE, echo=FALSE}
# Merge the dataset of electoral results with the dataset of controls

Spain_data <- merge(Spain_data_1, Controlvariables, by=c("Region","Year"), all=TRUE)
```

```{r adjustments, results='asis', tidy=TRUE, echo=FALSE}
#Final adjustments.  Change columns as numeric and clean one column

Spain_data$Unempl_rate_total<- as.numeric(Spain_data$Unempl_rate_total)
Spain_data$Unempl_rate_less25 <- as.numeric(Spain_data$Unempl_rate_less25)
Spain_data$Unempl_rate_16_19 <- as.numeric(Spain_data$Unempl_rate_16_19)
Spain_data$Unempl_rate_20_24 <- as.numeric(Spain_data$Unempl_rate_20_24)

Spain_data <- Spain_data[, -c(7,7)]
```


```{r fix, results='asis', tidy=TRUE, echo=FALSE}
# Create one dummy for each year
Spain_data["four"] <- as.numeric(Spain_data$Year == 2004)
Spain_data["eight"] <- as.numeric(Spain_data$Year == 2008)
Spain_data["eleven"] <- as.numeric(Spain_data$Year == 2011)
Spain_data["fifteen"] <- as.numeric(Spain_data$Year == 2015)
```

```{r fix, results='asis', tidy=TRUE, echo=FALSE}
# Exporting data
write.csv(Spain_data, "spain_data.csv")
```

I. INTRODUCTION AND RESEARCH QUESTION

Voter turnout is the most common measurement of participation in a democracy.Accordingly, the determinants of voter turnout is one the most widely studied topics by political scientists.

The turnout rate is influenced by many factors that may be related to age, race/ethnicity, gender, socio-economic status and political/economical situation of the country. In this broad spectrum of variables that influence participation of voters in elections, this article will focus specifically on whether youth unemployment (16 to 24 years old) has an impact on turnout in the case of Spain. Youth unemployment has been a pressing issue in the country in the period of 2000-2016, when the rates were not lower than 17.5\% throughout the whole period (Source: Eurostat).

We are also interested to see if youth unemployment mobilize voters with more intensity in regions where the rate of young people is higher. If this is the case, this effect can be explained by three different sources: because older people feel the resposability to vote more given the amount of young people unemployed, because young people go to vote more or because of a combination of the two factors above.

This case study is interesting for the recent change in the political system of the country, that went from a party system traditionally dominated by the PP (center-right) and the PSOE (center-left) to a multy party system where the Podemos (left) and Ciudadanos (liberal/center-right) reached the national parliament with a significant vote share in the recent national elections. Our guess is that, since youth unemployment rates in Spain was significantly high in recent years as a result of the euro crisis (reaching 48.1\%  in 2015), young voters turnout or other kind of voters infleunced by their feeling of responsability towards young people were among the reasons to explain why the party system changed so drastically in recent years.  

II. LITERATURE REVIEW

Literature on voter behavior tend to rely mainly on the incumbency-oriented hypothesis, which claims that voters tend to reward or punish government in elections according to the economic performance of the country (@norpoth1996presidents). In that sense, objective or subjective economic indicators such as unemployment rates, inflation, growth and other related aspects have the potential to explain part of the variance in government support (@lewis2000economic). Citizen dissatisfaction with economic performance substantially increases the possibility to vote against the incumbent. While some authors believe that the state of employment does not have a detectable effect upon voting behavior (@stigler1973general), others have reasons to believe that unemployment rates play a role depending on the political identification of the incumbent party, with left-wing governments suffering more from hign unemployment rates than right-wing governments (@dassonneville2013economic).

Another study goes beyond this polarization of the effect of unemployment on turnout by pointing out that individuals perceive job loss as a personal problem when the unemployment rate is low and a social problem when the unemployment rate is high. Therefore, job loss is a mobilizing experience when the unemployment rate is high. In particular, "unemployed Americans' political behavior is meaningfully influenced by unemployment context to an extent that we do not observe among gainfully employed Americans" (@incantalupo2011effects). This finding is particularly interesting in the context of our research. Although our study focuses on the Spanish case, we do not expect this kind of behavior to be completely different than the U.S case and this could mean that young unemployed, in a high unemployment context, would show higher turnout in the election day affecting the political outcome. 

Data from the Spanish Center of Sociological Research collected before the general elections in the Spring of 2011, shows that 40.6\% of the Spanish young people (15 to 29 years old) felt distrust regarding politics. It is relevant to mention that a few weeks after the general election occured in 2011, the 15-M movement started. This movement was  a "response to the inability of institutional politics to deal with the problems the country had been facing since 2009" and that speed up the appearance of new political parties into the National scene such as Podemos or Ciudadanos. In this context, @jover2014political argues that the previous distrust feeling would make young Spanish voters to not to participate in the general elections of 2011. Hence, after the 15-M and with the appearance of new political options, the youth would mobilize and vote more in subsequent elections. In fact, 41.8\% of the young people answered in a survey that the 15-M movement had influenced them a lot to cast their vote in the regional elections in the Fall of 2011, months after the General Elections. It is important to mention that @jover2014political shows data from self-reported surveys and does not prove causality between youth unemployment and turnout or between the appearance of new political parties and turnout.

III. STATISTICAL MODEL AND HYPOTHESIS

Our theory is that turnout is positively impacted by higher youth unemployment rates and even going further, that the effect of youth unemployment on tunout depends positively on the amount of young people. In that sense, we aim to reject the following hypotheses:

H0 = there is no relation between youth unemployment and turnout.

H1 = the amount of young people able to vote in a given region does not affect the relationship between youth unemployment and turnout.

In order to analyze this relation statistically, we propose the following model:

$Turnout_{i,t}=\beta_0+\beta_1 Youth\ Unemployment_{i,t}+\beta_2 \% Young\ People + \beta_3 Youth\ Unemployment*\% Young\ People + \delta_i + u_{i,t}$

where $i$ represents each of the 17 regions (Autonomus Communities) of Spain, $t$ represent the most recent years where general elections where held: 2004, 2008, 2011 and 2015 and $\delta_i$ represents regional fixed effects. 

Our model will be tested using data from the Spanish Statistic Office (INE) and the Electoral Results from the Spanish Ministry of Interior Affairs (MI) in the different regions of the country.



## Basic descriptive statistics

```{r 1, results='asis', tidy=TRUE, echo=FALSE, include=FALSE}
stat.desc(Spain_data) 
m <- stat.desc(Spain_data, basic=F)
m <- m[, -c(1:2)]
```


```{r stargazer, results='asis', tidy=TRUE, echo=FALSE, header = FALSE}
stargazer(m)
```

We are looking at tunout from general elections in Spain that occured in the years 2004, 2008, 2011 and 2015. The elections in 2015 did not result in the formation of government due to a highly fragmented parlament and the general elections were reperated in 2016. However, we decided to use only the data from the 2015 election since in a second call people might turn or not to turn to vote for different reasons than in normal elections.  The average turnout rate in Spain in these years (2004, 2008, 2011, 2015) was 72.02%. This is a pretty high turnout rate for a developed country that does not have compulsory voting. However, the differences across regions would be notable. Another thing to highlight from the table above is that invalid votes are in general more casted than blank votes.

When looking at unemployemnt rates it is clear that the average youth unemployment rate is much higher than the average total unemployment rate in years 2004, 2008, 2011 and 2015. Average youth unemployment rates increase drastically as the age of the populations groups studied lowers. Also, the standard deviation of the mean unemployment rate is higher for youth unemployemnt rate than for the normal unemployment rate indicating that youth unemployment varied more from year 2004 to year 2015.

##Turnout

```{r plot turnout, fig.width=8, fig.height=6, echo=FALSE}
ggplot(Spain_data, aes(x = Region, y = Turnout)) + geom_line()
```

The plot shows the variability of tunout by region in Spain. The regions that have lower turnout during the years 2004-2015 were Ceuta, Melilla, Baleares and Canarias. Ceuta and Melilla and autonomous cities that are phisically in Moroccan territory and Baleares and Canarias are islands. This gives the hint that people living in this regions might feel less attached to the Central government and therefore they vote less. It is also interesting to observe that the regions with a second official language different than Spanish, Catalunya, Galicia and Pais Vasco tend to have also lower turnout rates in the period.

##Unemployment

```{r plot unemployment, fig.width=8, fig.height=6, echo=FALSE}
ggplot(Spain_data, aes(x = Region, y = Unempl_rate_less25)) + geom_line()
```

Youth unemployment rates show a wide range during the period 2004-2015. Youth unemployment rates are structurally high in Spain. However, after the beginning of the economic crisis in 2008, youth unemployment rates explode. They went from around 20% in most regions to nearly 50%. However, Ceuta and Melilla stand out for their extremely high youth unemployment rates both before and after the economic crisis. The following graphs show that the youth unemployment rate is higher the lower is the age of the population studied.

```{r echo = FALSE, out.width=c('300px', '300px'), fig.show='hold'} 

```

```{r plot region 1, fig.width=8, fig.height=5, echo=FALSE, warning=FALSE}
# include map

region_code <- c("AN", "AR", "AS", "IB", "IC", "CA", "CM", "CL", "CT", "CE", "CV", "EX", "GA", "LR", "MA", "ME", "MU", "NV", "PV")

region_name <- c("Andalucia", "Aragon", "Asturias", "Baleares", "Canarias", "Cantabria", "Castilla_La_Mancha", "Castilla_y_Leon", "Catalunya", "Ceuta", "Comunitat_Valenciana", "Extremadura", "Galicia", "La_Rioja", "Madrid", "Melilla", "Murcia", "Navarra", "Pais_Vasco")

ggplot(data=Spain_data, aes(x=Year, y=Unempl_rate_20_24, group = Region, colour = Region)) + geom_line() + ggtitle("Youth Unemployment per region") + 
  labs(x="Year", y="Unemployment rate 20-24") + 
  ggtitle("Youth Unemployment per region 2004-2015") + 
  scale_color_discrete(labels = region_code) + 
  theme(legend.position = "bottom", legend.key = element_rect(size = 5, fill = 'white', colour = 'white'), legend.key.size = unit(2.0, 'lines'), axis.title.y=element_text(margin=margin(10,10,0,10)))

ggplot(data=Spain_data, aes(x=Year, y=Unempl_rate_20_24, group = Rich, colour = Region)) + geom_line() + ggtitle("Youth Unemployment per region") + 
  labs(x="Year", y="Unemployment rate 20-24") + 
  ggtitle("Youth Unemployment per region 2004-2015") + 
  scale_color_discrete(labels = region_code) + 
  theme(legend.position = "bottom", legend.key = element_rect(size = 5, fill = 'white', colour = 'white'), legend.key.size = unit(2.0, 'lines'), axis.title.y=element_text(margin=margin(10,10,0,10)))
```


```{r plot region 2, fig.width=8, fig.height=5, echo=FALSE, warning=FALSE}

plotmeans(Turnout ~ Region_code, main="Heterogeineity across regions", data=Spain_data)

# include map
```

## Unemployment and turnout


```{r unemp turn, fig.width=8, fig.height=6, echo=FALSE, warning=FALSE}
plot(Spain_data$Unempl_rate_total, Spain_data$turnout)
```

```{r unemp turn 1, fig.width=8, fig.height=6, echo=FALSE, warning=FALSE}
# Add aesthetic mappings

p <- ggplot(Spain_data, aes(Unempl_rate_less25, Turnout))
p + geom_point()
p1<-p + geom_point(aes(colour = Region_code)) 
p3 <- p1 + labs(y="Turnout", x= "Youth Unemployment")
p5 <- p3 + scale_y_continuous("Turnout",
                              limits=c(0.4,0.85),
                              breaks=seq(0.4, 0.85, 0.2)) + stat_smooth(method = 'lm', aes(colour = Region_code), se = FALSE)
p5

p6<-p5 + theme(axis.text.x = element_blank())
p6
```

# Unemployment and Turnout between Rich and Poor Regions

```{r unemp turn 2, fig.width=8, fig.height=6, echo=FALSE, warning=FALSE}
p <- ggplot(Spain_data, aes(Unempl_rate_less25, Turnout))
p + geom_point()
p1<-p + geom_point(aes(colour = Rich)) 
p3 <- p1 + labs(y="Turnout", x= "Youth Unemployment")
p5 <- p3 + scale_y_continuous("Turnout",
                              limits=c(0.4,0.85),
                              breaks=seq(0.4, 0.85, 0.2)) + stat_smooth(method = 'lm', aes(group = Rich), se = TRUE)
p5

Spain_data["Year_character"] <- as.character(Spain_data$Year)
```

# HIGHLIGHT SOME REGIONS WITH COLORS FOR YEARS
```{r unemp turn 3, fig.width=8, fig.height=6, echo=FALSE, warning=FALSE}
p <- ggplot(Spain_data, aes(Unempl_rate_less25, Turnout))
p + geom_point()
p1 <- p + geom_point(aes(colour = Year_character, group=Rich)) + 
  scale_colour_identity()
p3 <- p1 + labs(y="Turnout", x= "Youth Unemployment")
p5 <- p3 + scale_y_continuous("Turnout",
                              limits=c(0.4,0.85),
                              breaks=seq(0.4, 0.85, 0.2)) + stat_smooth(method = 'lm', aes(colour = Year_character, group=Rich), se = FALSE)
p5

p6<-p5 + theme(axis.text.x = element_blank())

p6

p7 <- p6 + ggtitle("Youth Unemployment per region") + 
  labs(x="Year", y="Unemployment rate 20-24") + 
  ggtitle("Youth Unemployment per region 2004-2015") + 
  scale_color_discrete(labels = region_code) + 
  theme(legend.position = "bottom", legend.key = element_rect(size = 5, fill = 'white', colour = 'white'), legend.key.size = unit(2.0, 'lines'), axis.title.y=element_text(margin=margin(10,10,0,10)))
p7
```

# LINEAR BETWEEN YEARS
```{r unemp turn 4, fig.width=8, fig.height=6, echo=FALSE, warning=FALSE}
p <- ggplot(Spain_data, aes(Unempl_rate_less25, Turnout))
p + geom_point()
p1 <- p + geom_point(aes(colour = Year_character)) + 
  scale_colour_identity()
p3 <- p1 + labs(y="Turnout", x= "Youth Unemployment")
p5 <- p3 + scale_y_continuous("Turnout",
                              limits=c(0.4,0.85),
                              breaks=seq(0.4, 0.85, 0.2)) + stat_smooth(method = 'lm', aes(colour = Year_character), se = FALSE)
p5

p6<-p5 + theme(axis.text.x = element_blank())

p6
``` 

# ORIGINAL - COLORS PER REGION AND LINES
```{r unemp turn 5, fig.width=8, fig.height=6, echo=FALSE, warning=FALSE}
p <- ggplot(Spain_data, aes(Unempl_rate_less25, Turnout))
p + geom_point()
p1<-p + geom_point(aes(colour = Region)) 
p3 <- p1 + labs(y="Turnout", x= "Youth Unemployment")
p5 <- p3 + scale_y_continuous("Turnout",
                              limits=c(0.4,0.85),
                              breaks=seq(0.4, 0.85, 0.2)) + stat_smooth(method = 'lm', aes(colour = Region), se = FALSE)
p5

p6<-p5 + theme(axis.text.x = element_blank())
p6
```

This plot shows how tunout and youth unemploment rate could be correlated. The plot shows that the correlation is not clear. 

The graphs are not showing the potential correlation by region that may exist between unemployment and turnout, an information that could give us a different perspective on the issue. Therefore, we need to explore this dataset further in order to conclude which is the relationship between turnout and youth unemployment rate in Spain.

##Bibliography