---
title: "Presentation"
author: "Rita Alvarez Martinez & Rafael Schmuziger"
date: "December 2nd, 2016"
output:
  beamer_presentation: default
  ioslides_presentation: default
  incremental: true

---

```{r options, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(message = FALSE)
# opts_chunk$set(tidy=FALSE)
# opts_chunk$set(fig.path='figs/', cache.path='cache/graphics-', 
#         fig.align='center', fig.width = 5, fig.height = 5, fig.show = 'hold',
#         cache = TRUE, par = TRUE)
```

```{r library, message=FALSE, echo=FALSE, include=FALSE}
library(pxR)
library(MicroDatosEs)
library(XML)
library(zoo)
library(rvest)
library(dplyr)
library(magrittr)
library(foreign)
library(readxl)
library(ggplot2)
library(pastecs)
library(stargazer)
library(gplots)
library(car)
```


```{r set wd, message=FALSE, echo=FALSE, include=FALSE}
setwd('C:/Users/Rita/Desktop/Final-Data-Project')
```

```{r turnout data, results='asis', tidy=TRUE, echo=FALSE}
# Read Turnout data

Turnout2004_2008_2011_2015 <- read_excel("C:/Users/Rita/Desktop/Assignment-3/ElectionsResult_2015_2011_2008_2004.xlsx", col_names = TRUE)
```

```{r unemployment data, results='asis', tidy=TRUE, echo=FALSE}
# Read Unemployment data

YouthUnemployment2004_2015 <- read_excel("C:/Users/Rita/Desktop/Assignment-3/YouthUnemployment2015_2011_2008_2004.xlsx", col_names = TRUE)
```

```{r dataset, results='asis', tidy=TRUE, echo=FALSE}
# Merge the dataset of electoral results with the dataset of unemployment rates. Change columns as numeric and clean one column

Spain_data <- merge(Turnout2004_2008_2011_2015, YouthUnemployment2004_2015, by=c("Region","Year"), all=TRUE)

Spain_data$Unempl_rate_total<- as.numeric(Spain_data$Unempl_rate_total)
Spain_data$Unempl_rate_less25 <- as.numeric(Spain_data$Unempl_rate_less25)
Spain_data$Unempl_rate_16_19 <- as.numeric(Spain_data$Unempl_rate_16_19)
Spain_data$Unempl_rate_20_24 <- as.numeric(Spain_data$Unempl_rate_20_24)

Spain_data <- Spain_data[, -c(7,7)]

Spain_data["Rich"] <- 1 

Spain_data$Rich <- ifelse(Spain_data$Region == 'Extremadura', 0, ifelse(Spain_data$Region == 'Castilla_la_Mancha', 0, ifelse(Spain_data$Region == 'Murcia', 0, ifelse(Spain_data$Region == 'Canarias', 0, ifelse(Spain_data$Region == 'Galicia', 0, ifelse(Spain_data$Region == 'Comunitat_Valenciana', 0, ifelse(Spain_data$Region == 'Melilla', 0, ifelse(Spain_data$Region == 'Asturias', 0, ifelse(Spain_data$Region == 'Ceuta', 0, ifelse(Spain_data$Region == 'Castilla_y_Leon', 0, ifelse(Spain_data$Region == 'Andalucia', 0, 1)))))))))))

Spain_data$Region_code <- ifelse(Spain_data$Region == 'Extremadura', 'EX', ifelse(Spain_data$Region == 'Castilla_La_Mancha', 'CM', ifelse(Spain_data$Region == 'Murcia', 'MU', ifelse(Spain_data$Region == 'Canarias', 'IC', ifelse(Spain_data$Region == 'Galicia', 'GA', ifelse(Spain_data$Region == 'Comunitat_Valenciana', 'CV', ifelse(Spain_data$Region == 'Melilla', 'ME', ifelse(Spain_data$Region == 'Asturias', 'AS', ifelse(Spain_data$Region == 'Ceuta', 'CE', ifelse(Spain_data$Region == 'Castilla_y_Leon', 'CL', ifelse(Spain_data$Region == 'Andalucia', 'AN', ifelse(Spain_data$Region == 'Cantabria', 'CA', ifelse(Spain_data$Region == 'Baleares', 'IB', ifelse(Spain_data$Region == 'Aragon', 'AR', ifelse(Spain_data$Region == 'Catalunya', 'CT', ifelse(Spain_data$Region == 'Madrid', 'MA', ifelse(Spain_data$Region == 'Navarra', 'NV', ifelse(Spain_data$Region == 'Pais_Vasco', 'PV', ifelse(Spain_data$Region == 'La_Rioja', 'LR', NA)))))))))))))))))))
```

#SUMMARY

- Motivation and Research Question

- Model

- Sources of data and methods of gathering

- Descriptive statistics
  
- Results

- Conclusion


#MOTIVATION

RECENT POLITICAL DEVELOPMENTS

- \textbf{Spanish political system:} From a bi-party system (PP and the PSOE) to a multy party system (PP, PSOE, Podemos and Ciudadanos) in less than two years. 

- \textbf{Critical social situation:} Especially for young people (youth unemployment rates reached 48.1%  in 2015).

- \textbf{Increased social mobilization:} The M15 movement (\textit{indignados}) and demonstrations against austerity measures mobilized people traditionally not involved with politics - especially young people.

#RESEARCH QUESTIONS

- \textbf{Literature:} Turnout rate related to age, race/ethnicity, gender, socio-economic status and political/economical factors. 

- Our model focus on whether \textbf{youth unemployment} (16 to 24 years old) has an impact on turnout in the case of Spain. 

\textbf{Do higher youth unemployment rates have impact on voter turnout?}  


#MODEL

Regional Fixed effects model
$$
(1)\ Turnout_{it}= \beta_{0}+\beta_{1}Youth\ Unemployment\ Rate_{it}+\delta_{i} + \epsilon_{it}
$$

Regional and Time Fixed effects model
$$
(2)\ Turnout_{it}= \beta_{0}+\beta_{1}Youth\ Unemployment\ Rate_{it}+\delta_{i} +\lambda_{t}+ \epsilon_{it}
$$

#SOURCES OF DATA

Dataset comprised by two sources: 

1. Electoral data of 2004, 2008, 2011 and 2015 (Spain Public Administration).
2. Unemployment rates by age groups in the same period (Spanish National Statistics Institute - INE).

#METHODS OF GATHERING

- \textbf{Electoral data:} \textbf{\textit{Web scraping}} we used a different URL source to grab the data available online for each of the 19 regions of Spain in the four election cycles.

- \textbf{Unemployment data:} We used the tool that the INE offers to download the needed data on Excel and imported from R.

Both datasets were cleaned/merged to form the final dataset "Spain_data".

#DESCRIPTIVE STATISTICS (1)

\textbf{Dependent Variable: Turnout}

```{r plot1,fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE}
boxplot(Turnout ~ Region_code, main="Heterogeineity across regions", data=Spain_data)
``` 

#DESCRIPTIVE STATISTICS (2)

\textbf{Dependent Variable: Turnout}

```{r plot2, fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE, warning=FALSE}
boxplot(Turnout ~ Year, main="Heterogeineity across years", data=Spain_data)
``` 

#DESCRIPTIVE STATISTICS (3)

\textbf{Independent Variable: Youth Unemployment}

```{r plot5,fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE}
d2004 <- subset(Spain_data, Year <= 2004)
d2008 <- subset(Spain_data, Year <= 2008 & Year>2004)
d2011 <- subset(Spain_data, Year <= 2011 & Year>2008)
d2015 <- subset(Spain_data, Year <= 20015 & Year>2011)

ggplot() +
geom_line(data=d2004, aes(x=Region_code, y=Unempl_rate_less25, group = 1, colour = "2004")) +
geom_line(data=d2008, aes(x=Region_code, y=Unempl_rate_less25, group = 1, colour = "2008")) +
geom_line(data=d2011, aes(x=Region_code, y=Unempl_rate_less25, group = 1, colour = "2011")) +
geom_line(data=d2015, aes(x=Region_code, y=Unempl_rate_less25, group = 1, colour = "2015")) + ggtitle("Youth Unemployment per region 2004-2015") + 
 labs(x="Region", y="Unemployment rate 16-24") + 
 ggtitle("Youth Unemployment per region 2004-2015") + 
 scale_color_discrete(labels = Spain_data$Year) + 
 theme(legend.position = "bottom", legend.key = element_rect(size = 5, fill = 'white', colour = 'white'), legend.key.size = unit(5.0, 'lines'), axis.title.y=element_text(margin=margin(10,10,0,10))) + theme(axis.title.x = element_text(size = 15, vjust=-.2)) +
 theme(axis.title.y = element_text(size = 15, vjust=0.3))
```

#DESCRIPTIVE STATISTICS (4)

Turnout and Youth Unemployment

```{r plot3,fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE}
p <- ggplot(Spain_data, aes(Unempl_rate_less25, Turnout))
p1<-p + geom_point(aes(colour = Region_code)) 
p3 <- p1 + labs(y="Turnout", x= "Youth Unemployment")
p5 <- p3 + scale_y_continuous("Turnout",
                              limits=c(0.4,0.85),
                              breaks=seq(0.4, 0.85, 0.2))
p6<-p5 + theme(axis.text.x = element_blank())
p6
``` 

#DESCRIPTIVE STATISTICS (5)

Turnout and Youth Unemployment patterns

```{r plot4,fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE}
p <- ggplot(Spain_data, aes(Unempl_rate_less25, Turnout))
p1<-p + geom_point(aes(colour = Region_code)) 
p3 <- p1 + labs(y="Turnout", x= "Youth Unemployment")
p5 <- p3 + scale_y_continuous("Turnout",
                              limits=c(0.4,0.85),
                              breaks=seq(0.4, 0.85, 0.2)) + stat_smooth(method = 'lm', aes(colour = Region_code), se = FALSE)
p6<-p5 + theme(axis.text.x = element_blank())
p6
``` 

#RESULTS (1)

\begin{table}[!htbp] \centering 
  \caption{Basic Regression} 
  \label{} 
\normalsize
\scalebox{0.55}
{
\begin{tabular}{@{\extracolsep{5pt}}lcc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
\cline{2-3} 
\\[-1.8ex] & \multicolumn{2}{c}{Voter Turnout in General Elections} \\ 
\\[-1.8ex] & (1) & (2)\\ 
\hline \\[-1.8ex] 
 Youth Unemloyment & $-$0.001$^{***}$ & $-$0.001$^{*}$ \\ 
  & (0.0002) & (0.001) \\ 
eight &  & $-$0.007 \\ 
  &  & (0.007) \\ 
eleven &  & $-$0.009 \\ 
  &  & (0.016) \\ 
 fifteen &  & 0.006 \\ 
  &  & (0.018) \\ 
 Constant & 0.771$^{***}$ & 0.777$^{***}$ \\ 
  & (0.013) & (0.021) \\ 
Region fixed effects & YES& YES \\
\hline \\[-1.8ex] 
Observations & 76 & 76 \\ 
R$^{2}$ & 0.925 & 0.933 \\ 
Adjusted R$^{2}$ & 0.899 & 0.906 \\ 
Residual Std. Error & 0.021 (df = 56) & 0.020 (df = 53) \\ 
F Statistic & 36.162$^{***}$ (df = 19; 56) & 33.701$^{***}$ (df = 22; 53) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular}
}
\end{table} 

#RESULTS (2)
```{r fix, results='asis', tidy=TRUE, echo=FALSE}
# Create one dummy for each year
Spain_data["four"] <- as.numeric(Spain_data$Year == 2004)
Spain_data["eight"] <- as.numeric(Spain_data$Year == 2008)
Spain_data["eleven"] <- as.numeric(Spain_data$Year == 2011)
Spain_data["fifteen"] <- as.numeric(Spain_data$Year == 2015)

Reg3 <- lm(Turnout ~ Unempl_rate_less25 + Region + eight + eleven + fifteen, data = Spain_data)
plot(Reg3,which=1)
```

#CONCLUSIONS

- In Spain, youth unemployment affects turnout negatively, in general.

\textbf{BUT:}

- Different regions present different patterns

- In 2015, a higher level of youth unemployment was correlated to higher tunout



#EXTRA SLIDE

```{r stargazer2, results='asis', tidy=TRUE, echo=FALSE, header = FALSE}
Reg2 <- lm(Turnout ~ Unempl_rate_less25 + Region, data = Spain_data)

# Create one dummy for each year
Spain_data["four"] <- as.numeric(Spain_data$Year == 2004)
Spain_data["eight"] <- as.numeric(Spain_data$Year == 2008)
Spain_data["eleven"] <- as.numeric(Spain_data$Year == 2011)
Spain_data["fifteen"] <- as.numeric(Spain_data$Year == 2015)

Reg3 <- lm(Turnout ~ Unempl_rate_less25 + Region + eight + eleven + fifteen, data = Spain_data)

stargazer(Reg2,Reg3,
          header=FALSE,
          title = "Basic Regression",
          covariate.labels = c("Youth Unemloyment", "Constant"),
          dep.var.labels   = "Voter Turnout in General Elections")
```