---
title: "Presentation"
author: "Rita Alvarez Martinez & Rafael Schmuziger"
date: "December 2nd, 2016"
output:
  beamer_presentation: default
  ioslides_presentation: default
  incremental: true
---

```{r options, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(message = FALSE)
# opts_chunk$set(tidy=FALSE)
# opts_chunk$set(fig.path='figs/', cache.path='cache/graphics-', 
#         fig.align='center', fig.width = 5, fig.height = 5, fig.show = 'hold',
#         cache = TRUE, par = TRUE)
```

```{r library, message=FALSE, echo=FALSE, include=FALSE}
library(pxR)
library(MicroDatosEs)
library(XML)
library(zoo)
library(rvest)
library(dplyr)
library(magrittr)
library(foreign)
library(readxl)
library(ggplot2)
library(pastecs)
library(stargazer)
library(gplots)
library(car)
```


```{r set wd, message=FALSE, echo=FALSE, include=FALSE}
setwd('C:/Users/Rita/Desktop/Final-Data-Project')
```

```{r turnout data, results='asis', tidy=TRUE, echo=FALSE}
# Read Turnout data

Turnout2004_2008_2011_2015 <- read_excel("C:/Users/Rita/Desktop/Assignment-3/ElectionsResult_2015_2011_2008_2004.xlsx", col_names = TRUE)
```

```{r unemployment data, results='asis', tidy=TRUE, echo=FALSE}
# Read Unemployment data

YouthUnemployment2004_2015 <- read_excel("C:/Users/Rita/Desktop/Assignment-3/YouthUnemployment2015_2011_2008_2004.xlsx", col_names = TRUE)
```

```{r dataset, results='asis', tidy=TRUE, echo=FALSE}
# Merge the dataset of electoral results with the dataset of unemployment rates. Change columns as numeric and clean one column

Spain_data <- merge(Turnout2004_2008_2011_2015, YouthUnemployment2004_2015, by=c("Region","Year"), all=TRUE)

Spain_data$Unempl_rate_total<- as.numeric(Spain_data$Unempl_rate_total)
Spain_data$Unempl_rate_less25 <- as.numeric(Spain_data$Unempl_rate_less25)
Spain_data$Unempl_rate_16_19 <- as.numeric(Spain_data$Unempl_rate_16_19)
Spain_data$Unempl_rate_20_24 <- as.numeric(Spain_data$Unempl_rate_20_24)

Spain_data <- Spain_data[, -c(7,7)]

Spain_data["Rich"] <- 1 

Spain_data$Rich <- ifelse(Spain_data$Region == 'Extremadura', 0, ifelse(Spain_data$Region == 'Castilla_la_Mancha', 0, ifelse(Spain_data$Region == 'Murcia', 0, ifelse(Spain_data$Region == 'Canarias', 0, ifelse(Spain_data$Region == 'Galicia', 0, ifelse(Spain_data$Region == 'Comunitat_Valenciana', 0, ifelse(Spain_data$Region == 'Melilla', 0, ifelse(Spain_data$Region == 'Asturias', 0, ifelse(Spain_data$Region == 'Ceuta', 0, ifelse(Spain_data$Region == 'Castilla_y_Leon', 0, ifelse(Spain_data$Region == 'Andalucia', 0, 1)))))))))))

Spain_data$Region_code <- ifelse(Spain_data$Region == 'Extremadura', 'EX', ifelse(Spain_data$Region == 'Castilla_La_Mancha', 'CM', ifelse(Spain_data$Region == 'Murcia', 'MU', ifelse(Spain_data$Region == 'Canarias', 'IC', ifelse(Spain_data$Region == 'Galicia', 'GA', ifelse(Spain_data$Region == 'Comunitat_Valenciana', 'CV', ifelse(Spain_data$Region == 'Melilla', 'ME', ifelse(Spain_data$Region == 'Asturias', 'AS', ifelse(Spain_data$Region == 'Ceuta', 'CE', ifelse(Spain_data$Region == 'Castilla_y_Leon', 'CL', ifelse(Spain_data$Region == 'Andalucia', 'AN', ifelse(Spain_data$Region == 'Cantabria', 'CA', ifelse(Spain_data$Region == 'Baleares', 'IB', ifelse(Spain_data$Region == 'Aragon', 'AR', ifelse(Spain_data$Region == 'Catalunya', 'CT', ifelse(Spain_data$Region == 'Madrid', 'MA', ifelse(Spain_data$Region == 'Navarra', 'NV', ifelse(Spain_data$Region == 'Pais_Vasco', 'PV', ifelse(Spain_data$Region == 'La_Rioja', 'LR', NA)))))))))))))))))))
```

#SUMMARY

- Motivation and Research Question

- Model

- Sources of data and methods of gathering

- Descriptive statistics
  
- Results

- Conclusion


#MOTIVATION

RECENT POLITICAL DEVELOPMENTS

- \textbf{Spanish political system:} From a bi-party system (PP and the PSOE) to a multy party system (PP, PSOE, Podemos and Ciudadanos) in less than two years. 

- \textbf{Critical social situation:} Especially for young people (youth unemployment rates reached 48.1%  in 2015).

- \textbf{Increased social mobilization:} The M15 movement (\textit{indigandos}) and demonstrations against austerity measures mobilized people traditionally not involved with politics - especially young people.

#RESEARCH QUESTIONS

- \textbf{Literature:} Turnout rate related to age, race/ethnicity, gender, socio-economic status and political/economical factors. 

- Our model focus on whether \textbf{youth unemployment} (16 to 24 years old) has an impact on turnout in the case of Spain. 

\textbf{Do higher youth unemployment rates have impact on voter turnout?}  


#MODEL

Regional Fixed effects model
$$
Turnout_{it}= \beta_{0}+\beta_{1}Youth\ Unemployment\ Rate_{it}+\delta_{i} + \epsilon_{it}
$$

Regional and Time Fixed effects model
$$
Turnout_{it}= \beta_{0}+\beta_{1}Youth\ Unemployment\ Rate_{it}+\delta_{i} +\lambda_{t}+ \epsilon_{it}
$$

#SOURCES OF DATA

Dataset comprised by two sources: 

1. Electoral data of 2004, 2008, 2011 and 2015 (Spain Public Administration).
2. Unemployment rates by age groups in the same period (Spanish National Statistics Institute - INE).

#METHODS OF GATHERING

- \textbf{Electoral data:} \textbf{\textit{Web scraping}} we used a different URL source to grab the data available online for each of the 19 regions of Spain in the four election cycles.

- \textbf{Unemployment data:} We used the tool that the INE offers to download the needed data on Excel and imported from R.

Both datasets were cleaned/merged to form the final dataset "Spain_data".

#DESCRIPTIVE STATISTICS (1)

\textbf{Dependent Variable: Turnout}

```{r plot1,fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE}
boxplot(Turnout ~ Region_code, main="Heterogeineity across regions", data=Spain_data)
``` 

#DESCRIPTIVE STATISTICS (2)

\textbf{Dependent Variable: Turnout}

```{r plot2, fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE, warning=FALSE}
boxplot(Turnout ~ Year, main="Heterogeineity across years", data=Spain_data)
``` 

#DESCRIPTIVE STATISTICS (3)

\textbf{Independent Variable: Youth Unemployment}

```{r plot5,fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE}
d2004 <- subset(Spain_data, Year <= 2004)
d2008 <- subset(Spain_data, Year <= 2008 & Year>2004)
d2011 <- subset(Spain_data, Year <= 2011 & Year>2008)
d2015 <- subset(Spain_data, Year <= 20015 & Year>2011)

ggplot() +
geom_line(data=d2004, aes(x=Region_code, y=Unempl_rate_less25, group = 1), color='black') +
geom_line(data=d2008, aes(x=Region_code, y=Unempl_rate_less25, group = 1), color='red') +
geom_line(data=d2011, aes(x=Region_code, y=Unempl_rate_less25, group = 1), color='green') +
geom_line(data=d2015, aes(x=Region_code, y=Unempl_rate_less25, group = 1), color='blue')
```

#DESCRIPTIVE STATISTICS (4)

Turnout and Youth Unemployment

```{r plot3,fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE}
p <- ggplot(Spain_data, aes(Unempl_rate_less25, Turnout))
p1<-p + geom_point(aes(colour = Region_code)) 
p3 <- p1 + labs(y="Turnout", x= "Youth Unemployment")
p5 <- p3 + scale_y_continuous("Turnout",
                              limits=c(0.4,0.85),
                              breaks=seq(0.4, 0.85, 0.2))
p6<-p5 + theme(axis.text.x = element_blank())
p6
``` 

#DESCRIPTIVE STATISTICS (5)

Turnout and Youth Unemployment patterns

```{r plot4,fig.width=10, fig.height=6, results='asis', tidy=TRUE, echo=FALSE}
p <- ggplot(Spain_data, aes(Unempl_rate_less25, Turnout))
p1<-p + geom_point(aes(colour = Region_code)) 
p3 <- p1 + labs(y="Turnout", x= "Youth Unemployment")
p5 <- p3 + scale_y_continuous("Turnout",
                              limits=c(0.4,0.85),
                              breaks=seq(0.4, 0.85, 0.2)) + stat_smooth(method = 'lm', aes(colour = Region_code), se = FALSE)
p6<-p5 + theme(axis.text.x = element_blank())
p6
``` 

#RESULTS

```{r stargazer2, results='asis', tidy=TRUE, echo=FALSE, header = FALSE}
Reg2 <- lm(Turnout ~ Unempl_rate_less25 + Region, data = Spain_data)
stargazer(Reg2)
```

#CONCLUSIONS